---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

# openfda-extract

<!-- badges: start -->
<!-- badges: end -->

This repository collects [adverse events data](https://open.fda.gov/apis/device/event/download/) from openFDA. A single [script](/data-raw/loop.R) attempts to:

1. Convert JSON files to tabular format using `jsonlite::fromJSON` and `tidyr::unnest`.
2. Save the data to a database.

The data is relational as described [here](https://opendata.stackexchange.com/a/2187). Converting the data to tabular format may not be efficient and causes lots of duplication. To avoid duplication, I have tried to store nested data in different tables:

1. `adverse_events`
2. `adverse_events.mdr_text`
3. `adverse_events.product_problems`
4. `adverse_events.source_type`
5. `adverse_events.device`
6. `adverse_events.patient`
7. `adverse_events.remedial_action`
8. `adverse_events.type_of_report`

Where the naming convention is `mainframe.<list col>`.

## Hardware

1. I transformed the data on:
    * 2013 15" MacBook Pro, 8 GB Memory, 8 Core CPU.
2. I wrote the data to:
    * Postgres database hosted on a digital ocean droplet
    * 2 GB Memory, 2 vCPUs, 60 GB Disk, Ubuntu 18.04.3 (LTS) x64
    
For context, this was the result of my first run:

```
~ openFDA database refresh completed in [17.4161201466454 hours]
```

## Examples

If you have successfully ran everything, you should have 8 tables with millions of observations that you can explore.

```{r}
library(dplyr, warn.conflicts = FALSE)
library(DBI)

# credentials
dw <- config::get("datawarehouse")

# connect to db
con <- DBI::dbConnect(
  odbc::odbc(),
  Driver = dw$driver,
  Server = dw$server,
  Database = dw$database,
  UID = dw$uid,
  PWD = dw$pwd,
  Port = dw$port
)

# list all available tables
dbListTables(con)

# query the mdr text
tbl(con, "adverse_events.mdr_text") %>% 
  select(id, text)

# query the device information
tbl(con, "adverse_events.device") %>% 
  select(id, manufacturer_d_name)

# disconnect
dbDisconnect(con)
```

Note that there is an `id` column in every table, for example:

* `2014q4-0002-0003-3683`
* `<year quarter>-<part n>-<n parts>-<row number>`

I made this column so that tables can be joined (though this ID in some other form might already exist in the data, I just haven't figured out if that is the case or not).
